---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # fig.path = "man/figures/README-",
  fig.path = "README-",
  out.width = "100%"
)
```

# EDIutils

<!-- badges: start -->
[![codecov.io](https://codecov.io/github/EDIorg/EDIutils/coverage.svg?branch=master)](https://codecov.io/github/EDIorg/EDIutils?branch=master)
<!-- badges: end -->

## Overview

A client for the Environmental Data Initiative (EDI) repository. Provides create, read, update, and delete operations to the EDI Data Package and Audit Manager APIs. The EDI Data Repository (https://portal.edirepository.org/nis/home.jsp) is an open access repository for anyone curating or reusing high quality environmental and ecological data.

## Installation

Get the latest development version:
```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("EDIorg/EDIutils")
```

## Usage

TODO

## Getting help

Use [GitHub Issues](https://github.com/EDIorg/EDIutils/issues) for bug reporting, feature requests, and general questions/discussions. When filing bug reports, please include a minimal reproducible example.

## Contributing

Community contributions are welcome! Please reference our [contributing guidelines](https://github.com/EDIorg/EDIutils/blob/master/CONTRIBUTING.md) for details.

-----

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/EDIorg/EDIutils/blob/master/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.