---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # fig.path = "man/figures/README-",
  fig.path = "README-",
  out.width = "100%"
)
```

# EDIutils

<!-- badges: start -->
[![codecov.io](https://codecov.io/github/EDIorg/EDIutils/coverage.svg?branch=master)](https://codecov.io/github/EDIorg/EDIutils?branch=master)
<!-- badges: end -->

A client for the Environmental Data Initiative (EDI) repository (https://portal.edirepository.org/nis/home.jsp). Provides create, read, update, and delete operations for the Data Package Manager and Audit Manager APIs. The EDI Data Repository is an open access repository for anyone curating or reusing high quality environmental and ecological data.

[Overview of the EDI repository]()

## Installation

Get the latest development version:
```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("EDIorg/EDIutils")
```

## Usage

[Access data]()
[Upload data]()
[Report to stakeholders]()
[Analyze holdings]()

## Getting help

Use [GitHub Issues](https://github.com/EDIorg/EDIutils/issues) for bug reporting, feature requests, and general questions/discussions. When filing bug reports, please include a minimal reproducible example.

## Contributing

Community contributions are welcome! Please reference our [contributing guidelines](https://github.com/EDIorg/EDIutils/blob/master/CONTRIBUTING.md) for details.

-----

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/EDIorg/EDIutils/blob/master/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.