---
title: "Evaluate and Upload Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Evaluate and Upload Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EDIutils)
```

### Evaluate and Upload Data

_The EDI repository has a "staging" environment to test the upload and rendering of new data packages before publishing in the "production" environment._

Evaluation and upload to the EDI repository requires data entities are described with EML metadata. There are many tools for creating EML, EDI supports two: the [EMLassemblyline](https://ediorg.github.io/EMLassemblyline/) R package for programmatic workflows, and the [ezEML](https://ezeml.edirepository.org/eml/de) web form wizard for interactive metadata generation. Research groups managing large volumes of metadata should consider the [LTER core-metabase]().

##### Authenticate

Authentication is required by functions involving data evaluation and upload, audit report access, event notifications, and other account based features. Request an account with support@environmentaldatainitiative.org. There are three options for authenticating:

```{r eval=FALSE}
# Interactively at the console
login()
#> User name: "my_name"
#> User password: "my_secret"

# Programmatically with function arguments
login(userId = "my_name", userPass = "my_secret")

# Programmatically with a file containing userId and userPass arguments
login(config = "./data/config.txt")
```

This function exchanges your credentials for a temporary authentication token (~10 hour) is written to the `EDI_USERID` system variable referenced by EDIutils functions.

##### Reserve Data Package ID

Data package reservations prevent conflicting use of the same identifier.

```{r eval=FALSE}
# Create reservation
identifier <- create_reservation(scope = "edi", env = "staging")
identifier
#> [1] 595
```

#### Evaluate

Evaluation checks metadata accuracy and completeness.

```{r eval=FALSE}

# Evaluate data package
transaction <- evaluate_data_package(
 eml = "./data/edi.595.1.xml", 
 env = "staging")
transaction
#> [1] "evaluate_163966785813042760"

# Check status
status <- check_status_evaluate(transaction, env = "staging")
status
#> [1] TRUE

# Read the evaluation report
report <- read_evaluate_report(transaction, frmt = "char", env = "staging")
report
```


#### Upload

Upload once errors and warnings are fixed.

```{r eval=FALSE}
# Create a new data package
transaction <- create_data_package(
 eml = "./data/edi.595.1.xml", 
 env = "staging")
transaction
#> [1] "create_163966765080210573__edi.595.1"

# Check status
status <- check_status_create(
 transaction = transaction, 
 env = "staging")
status
#> [1] TRUE
```

#### Update

Update is the same as upload, but with an incremented data package version number (e.g. "edi.595.1" is superseded by "edi.595.2").

```{r eval=FALSE}
#' # Update data package
#' transaction <- update_data_package(
#'   eml = "./data/edi.595.2.xml", 
#'   env = "staging")
#' transaction
#' #> [1] "update_edi.595_163966788658131920__edi.595.2"
#' 
#' # Check status
#' status <- check_status_update(
#'   transaction = transaction, 
#'   env = "staging")
#' status
#' #> [1] TRUE
```

Once everything looks good in the "staging" environment, then repeat the above reservation and upload steps in the "production" environment where it will be assigned a DOI and officially published.
