---
title: "Tests Requiring Authentication"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tests Requiring Authentication}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

_This vignette is for `EDIutils` maintainers and developers._

Some tests require authentication and a minimal data package for evaluation and upload. This vignette demonstrates how to create and run these tests with `testthat`.

```{r setup}
library(EDIutils)
```

## Get an EDI User Account

Request accounts via info@environmentaldatainitiative.org

## Use the "staging" Environment!
The EDI repository "staging" environment is a sandbox for testing data package rendering, etc. Do not use the "production" environment for testing. The "production" environment is where publication quality data are released.

## Create Tests

#### Requiring Authentication

The first line of any test requiring authentication should include `skip_if_logged_out()`. This is an internal use only function located in R/utilities.R. Tests including this line will be skipped unless the R environment variable `EDI_TOKEN` is set. Use `login()` to set this variable.

Example:
```{r eval=FALSE}
context("Create journal citation")

testthat::test_that("Test attributes of returned object", {
  skip_if_logged_out()
  journalCitationId <- create_journal_citation(
    packageId = get_test_package(), 
    articleDoi = "10.1890/11-1026.1",
    articleUrl = "https://doi.org/10.1890/11-1026.1",
    articleTitle = "Corridors promote fire via connectivity and edge effects",
    journalTitle = "Ecological Applications",
    relationType = "IsCitedBy",
    env = "staging")
  expect_type(journalCitationId, "double")
  res <- delete_journal_citation(journalCitationId, env = "staging")
  expect_true(res)
})
```

#### Requiring Data Package Evaluation or Upload

If evaluation or upload of a data package is required by the test, then include a second line `skip_if_missing_eml_config()`. Tests including this line will be skipped unless the R environment variables `EDI_USERID` and `EDI_TEST_URL` are set. Use `config_test_eml()` to set these variables.

To set up the minimal test data package:

1. Copy the test data object at inst/extdata/data.txt to a web accessible location. The EDI repository will download this data object by it's URL. The URL cannot present any redirects or else the EDI repository will not have access.

2. Add `create_test_eml()` to create an EML metadata file within the context of the test. Data packages require unique identifiers that often change from test to test.

Example:
```{r eval=FALSE}
context("Evaluate data package")

testthat::test_that("Test attributes of returned object", {
  skip_if_logged_out()
  skip_if_missing_eml_config()
  # Create data package for evaluation
  identifier <- create_reservation(scope = "edi", env = "staging")
  packageId <- paste0("edi.", identifier, ".1")
  eml <- create_test_eml(path = tempdir(), packageId = packageId)
  on.exit(file.remove(eml), add = TRUE, after = FALSE)
  # Evaluate
  transaction <- evaluate_data_package(eml, env = "staging")
  res <- check_status_evaluate(transaction, env = "staging")
  expect_true(res)
  # Read evaluation report
  report <- read_evaluate_report(transaction, env = "staging")
  expect_true("xml_document" %in% class(report))
  delete_reservation("edi", identifier, env = "staging")
})

```

## Run Tests

1. Authenticate with `login()`

2. Configure test data package with `config_test_eml()`

3. Run test suite `devtools::test()`
